---
title: "How to Make High Quality Data Visualizations for Websites With R and ggplot2"
author: "Max Woolf (@minimaxir)"
date: "2014-08-14"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: yes
    theme: spacelab
    toc: yes
---


# Setup

```{r}
library(tidyverse)
library(lubridate)
library(tidytext)
library(scales)
library(viridis)

sessionInfo()

# https://brandcolors.net/b/stackoverflow
stack_overflow_color <- "#f48024"
```

```{r}
theme_set(theme_minimal(base_size=9, base_family="Source Sans Pro") +
            theme(plot.title = element_text(size=8, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.1, unit='cm')),
                  axis.title.x = element_text(size=8),
                  axis.title.y = element_text(size=8),
                  plot.subtitle = element_text(family="Source Sans Pro Semibold", color="#969696", size=6),
                  plot.caption = element_text(size=6, color="#969696"),
                  #axis.text.x = element_text(size = 5),
                  legend.text = element_text(size = 6),
                  legend.key.width = unit(0.25, unit='cm')))
```


```{r}
file_path <- "~/Downloads/stack-overflow-2017.csv"

df <- read_csv(file_path)

df %>% head()
```

Add columns relevant to the timing when the post was made. The raw data is in `UTC`, so it must be converted to Eastern.

```{r}
df <- df %>% mutate(
        creation_date = with_tz(creation_date, "America/New_York"),
        hour_posted = hour(creation_date),
        weekday_posted = wday(creation_date, label=T, abbr=F),
        week_posted = floor_date(creation_date, '1 week'),
        month_posted = floor_date(creation_date, '1 month')
)
```

Add a few bespoke features:

```{r}
df <- df %>% mutate(
        f_answer_in = ifelse(!is.na(time_to_f) & time_to_f < 60*60*8, 1, 0),
        a_answer_in = ifelse(!is.na(time_to_a) & time_to_a < 60*60*8, 1, 0)
)
```


# Overview

For all questions asked:

## Weekly

```{r}
plot <- ggplot(df, aes(x=week_posted, y=..count..)) +
          geom_bar(fill = stack_overflow_color, alpha=0.75) +
          scale_x_datetime(date_breaks='2 months', date_labels='%b') +
          scale_y_continuous(labels=comma) +
          labs(title='New Stack Overflow Questions in 2017',
               x='Week Question was Posted',
               y='# of Questions Posted')

ggsave('weekly_count.png', plot, width=4, height=3)
```

# Facet by Tags

```{r}
df_top_tags <- df %>%
                select(creation_date, tags) %>%
                unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
                group_by(tag) %>%
                summarize(count=n()) %>%
                arrange(desc(count)) %>%
                head(1000)

df_top_tags %>% head()               
```

```{r}
df_top_tags_month <- df %>%
                select(month_posted, tags) %>%
                unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
                filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
                group_by(month_posted, tag) %>%
                summarize(count=n()) %>%
                arrange(month_posted) %>%
                filter(month_posted >= date('2017-01-01'))

df_top_tags_month %>% head()               
```

## Monthly by Tags

```{r}
plot <- ggplot(df_top_tags_month, aes(x=month_posted, y=count, fill=tag)) +
          geom_bar(alpha=0.75, stat="identity") +
          geom_smooth(se=F, method="lm", color="black", size=0.25) +
          scale_x_datetime(date_breaks='3 months', date_labels='%b') +
          scale_y_continuous(labels=comma) +
          facet_wrap(~ tag, nrow=10, ncol=4, scales="free_y") +
          labs(title='New Stack Overflow Questions for the Top 40 Tags',
                subtitle='From January 2017 to November 2017',
               x='Month Question Was Posted',
               y='# of Questions Posted With Tag') +
          theme(legend.position = 'none',
                plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')))

ggsave('monthly_count_tag.png', plot, width=6, height=8)
```

## Day/Hour of Week tags were posted

Related: https://stackoverflow.blog/2017/04/19/programming-languages-used-late-night/

```{r}
df_top_tags_hr_doy <- df %>%
                select(hour_posted, weekday_posted, tags) %>%
                unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
                filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
                group_by(hour_posted, weekday_posted, tag) %>%
                summarize(count=n()) %>%
                ungroup() %>%
                group_by(tag) %>%
                mutate(proportion=count/sum(count))

df_top_tags_hr_doy %>% head() 
```

```{r}
plot <- ggplot(df_top_tags_hr_doy, aes(x=hour_posted, y=fct_rev(weekday_posted), fill=proportion)) +
          geom_raster(stat="identity", interpolate=F) +
          scale_x_discrete() +
          scale_y_discrete() +
          scale_fill_viridis(option="inferno", labels=percent) +
          facet_wrap(~ tag, nrow=10, ncol=4) +
          labs(title='New Stack Overflow Questions for the Top 40 Tags, by Time Posted',
                subtitle='From January 2017 to November 2017',
               x='Hour Question Was Posted (12 AM - 11 PM Eastern Time)',
               y='Day of Week Question Was Posted',
               fill='Proportion of All Questions\nPosted w/ Tag',
               caption = "Max Woolf — minimaxir.com") +
          theme(legend.position = 'top',
                plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
                axis.text.y = element_text(size = 5),
                legend.text = element_text(size = 6),
                legend.title = element_text(size = 6),
                  legend.key.width = unit(1, unit='cm'),
                legend.key.height = unit(0.25, unit='cm'),
                legend.margin = margin(c(0, 0, -0.5, 0), unit='cm'))

ggsave('monthly_count_hr_doy.png', plot, width=6, height=8)
```

## Day/Week To Receive Answer


```{r}
df_top_tags_hr_doy_answer <- df %>%
                select(hour_posted, weekday_posted, tags, f_answer_in, a_answer_in) %>%
                unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
                filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
                group_by(hour_posted, weekday_posted, tag) %>%
                summarize(count=n(), perc_f_answer = sum(f_answer_in)/count, perc_a_answer = sum(a_answer_in)/count) %>%
                ungroup()

df_top_tags_hr_doy_answer %>% head() 
```

```{r}
plot <- ggplot(df_top_tags_hr_doy_answer, aes(x=hour_posted, y=fct_rev(weekday_posted), fill=perc_f_answer)) +
          geom_raster(stat="identity", interpolate=F) +
          scale_x_discrete() +
          scale_y_discrete() +
          scale_fill_distiller(palette="PRGn", labels=percent, limits=c(0,1)) +
          facet_wrap(~ tag, nrow=10, ncol=4) +
          labs(title='New Stack Overflow Questions for the Top 40 Tags, by Time Posted',
                subtitle='From January 2017 to November 2017',
               x='Hour Question Was Posted (12 AM - 11 PM Eastern Time)',
               y='Day of Week Question Was Posted',
               fill='Proportion of All Questions\nPosted w/ Tag',
               caption = "Max Woolf — minimaxir.com") +
          theme(legend.position = 'top',
                plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
                axis.text.y = element_text(size = 5),
                legend.text = element_text(size = 6),
                legend.title = element_text(size = 6),
                  legend.key.width = unit(1, unit='cm'),
                legend.key.height = unit(0.25, unit='cm'),
                legend.margin = margin(c(0, 0, -0.5, 0), unit='cm'))

ggsave('monthly_count_hr_doy_f_answer.png', plot, width=6, height=8)
```


```{r}
plot <- ggplot(df_top_tags_hr_doy_answer, aes(x=hour_posted, y=fct_rev(weekday_posted), fill=perc_a_answer)) +
          geom_raster(stat="identity", interpolate=F) +
          scale_x_discrete() +
          scale_y_discrete() +
          scale_fill_distiller(palette="PRGn", labels=percent, limits=c(0,1)) +
          facet_wrap(~ tag, nrow=10, ncol=4) +
          labs(title='New Stack Overflow Questions for the Top 40 Tags, by Time Posted',
                subtitle='From January 2017 to November 2017',
               x='Hour Question Was Posted (12 AM - 11 PM Eastern Time)',
               y='Day of Week Question Was Posted',
               fill='Proportion of All Questions\nPosted w/ Tag',
               caption = "Max Woolf — minimaxir.com") +
          theme(legend.position = 'top',
                plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
                axis.text.y = element_text(size = 5),
                legend.text = element_text(size = 6),
                legend.title = element_text(size = 6),
                  legend.key.width = unit(1, unit='cm'),
                legend.key.height = unit(0.25, unit='cm'),
                legend.margin = margin(c(0, 0, -0.5, 0), unit='cm'))

ggsave('monthly_count_hr_doy_a_answer.png', plot, width=6, height=8)
```
# LICENSE

The MIT License (MIT)

Copyright (c) 2017 Max Woolf

Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the "Software"), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.