---
title: "How to Make High Quality Data Visualizations for Websites With R and ggplot2"
author: "Max Woolf (@minimaxir)"
date: "2014-08-14"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: yes
    theme: spacelab
    toc: yes
---


# Setup

```{r}
library(tidyverse)
library(lubridate)
library(tidytext)
library(scales)
library(viridis)
library(ggrepel)

sessionInfo()

# https://brandcolors.net/b/stackoverflow
stack_overflow_color <- "#f48024"
```

```{r}
theme_set(theme_minimal(base_size=9, base_family="Source Sans Pro") +
            theme(plot.title = element_text(size=8, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.1, unit='cm')),
                  axis.title.x = element_text(size=8),
                  axis.title.y = element_text(size=8),
                  plot.subtitle = element_text(family="Source Sans Pro Semibold", color="#969696", size=6),
                  plot.caption = element_text(size=6, color="#969696"),
                  #axis.text.x = element_text(size = 5),
                  legend.text = element_text(size = 6),
                  legend.key.width = unit(0.25, unit='cm')))
```

# Behavior for new submissions

Use data precomputed from this BigQuery:

```{sql eval=FALSE, include=FALSE}
#standardSQL
SELECT
  DATE_TRUNC(DATE(creation_date), YEAR) AS year,
  SUM(view_count_delta) AS total_delta 
FROM (
  SELECT
    id,
    creation_date,
    b.view_count - a.view_count AS view_count_delta
  FROM
    `fh-bigquery.stackoverflow_archive.201703_posts_questions` a
  LEFT JOIN (
    SELECT
      id,
      view_count
    FROM
      `fh-bigquery.stackoverflow_archive.201712_posts_questions` ) b
  USING
    (id) )
GROUP BY
  year
ORDER BY
  year ASC

```


```{r}
file_path <- "stack_overflow_delta.csv"
df_deltas <- read_csv(file_path) %>% mutate(perc = total_delta / sum(as.numeric(total_delta)))
df_deltas %>% head()
```


```{r}
plot <- ggplot(df_deltas %>% filter(year >= ymd('2009-01-01'), year <= ymd('2016-01-01')), aes(x=year, y=perc)) +
          geom_bar(alpha=0.75, stat="identity", fill=stack_overflow_color) +
          scale_x_date(date_breaks='1 year', date_labels='%Y') +
          scale_y_continuous(labels=percent) +
          labs(title='Proportion of 2017 Views on Older Stack Overflow Questions by Year',
                subtitle='From March 13th, 2017 to December 3rd, 2017. Visualization Excludes Partial Years',
               x='Year Question Was Posted',
               y='% of All Views')

ggsave('so_overview.png', plot, width=4, height=3)
```

# Main Analysis

Data processed from this BigQuery:

```{sql eval=FALSE, include=FALSE}
#standardSQL
WITH
  answers_ordered AS (
  SELECT
    id,
    creation_date,
    parent_id AS question_id,
    score,
    ROW_NUMBER() OVER (PARTITION BY parent_id ORDER BY creation_date ASC) AS time_rank,
    ROW_NUMBER() OVER (PARTITION BY parent_id ORDER BY score DESC) AS score_rank,
    COUNT(*) OVER (PARTITION BY parent_id) AS num_answers
  FROM
    `fh-bigquery.stackoverflow_archive.201712_posts_answers` )
SELECT
  id,
  title,
  tags,
  DATETIME(creation_date) AS creation_date,
  accepted_answer_id,
  view_count,
  score,
  num_answers,
  f_answer_id,
  DATETIME(f_creation_date) AS f_creation_date,
  f_score,
  f_score_rank,
  a_answer_id,
  DATETIME(a_creation_date) AS a_creation_date,
  a_score,
  a_time_rank,
  a_score_rank,
  TIMESTAMP_DIFF(f_creation_date, creation_date, SECOND) AS time_to_f,
  TIMESTAMP_DIFF(a_creation_date, creation_date, SECOND) AS time_to_a
FROM
  `fh-bigquery.stackoverflow_archive.201712_posts_questions` q
LEFT JOIN (
  SELECT
    id AS f_answer_id,
    creation_date AS f_creation_date,
    question_id AS f_question_id,
    score AS f_score,
    score_rank AS f_score_rank,
    num_answers
  FROM
    answers_ordered
  WHERE
    time_rank = 1 ) f
ON
  q.id = f.f_question_id
LEFT JOIN (
  SELECT
    id AS a_answer_id,
    creation_date AS a_creation_date,
    score AS a_score,
    time_rank AS a_time_rank,
    score_rank AS a_score_rank
  FROM
    answers_ordered ) a
ON
  q.accepted_answer_id = a.a_answer_id
WHERE
  creation_date >= '2017-01-01 00:00:00' AND creation_date < '2017-12-01 00:00:00'
```


```{r}
file_path <- "~/Downloads/stack_overflow_2017.csv"
df <- read_csv(file_path)
df %>% head()
```

Add columns relevant to the timing when the post was made. The raw data is in `UTC`, so it must be converted to Eastern.

```{r}
df <- df %>% mutate(
  creation_date = with_tz(creation_date, "America/New_York"),
  hour_posted = hour(creation_date),
  weekday_posted = wday(creation_date, label=T, abbr=F),
  week_posted = floor_date(creation_date, '1 week'),
  month_posted = floor_date(creation_date, '1 month')
)
```

Add a few bespoke features:

```{r}
df <- df %>% mutate(
  num_answers = ifelse(is.na(num_answers), 0, num_answers),
  f_answer_in = ifelse(!is.na(time_to_f) & time_to_f < 60*60*4, 1, 0),
  a_answer_in = ifelse(!is.na(time_to_a) & time_to_a < 60*60*4, 1, 0)
)
```


# Overview

For all questions asked:

## Weekly

```{r}
plot <- ggplot(df, aes(x=week_posted, y=..count..)) +
  geom_bar(fill = stack_overflow_color, alpha=0.75) +
  scale_x_datetime(date_breaks='2 months', date_labels='%b') +
  scale_y_continuous(labels=comma) +
  labs(title='New Stack Overflow Questions in 2017',
       x='Week Question was Posted',
       y='# of Questions Posted')

ggsave('weekly_count.png', plot, width=4, height=3)
```

# Facet by Tags

```{r}
df_top_tags <- df %>%
  select(creation_date, tags, score, num_answers) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  group_by(tag) %>%
  summarize(count = n(),
            med_score = median(score),
            med_answers = median(num_answers)) %>%
  arrange(desc(count)) %>%
  head(1000)

df_top_tags %>% head()               
```

```{r}
df_top_tags_month <- df %>%
  select(month_posted, tags) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
  group_by(month_posted, tag) %>%
  summarize(count=n()) %>%
  arrange(month_posted) %>%
  filter(month_posted >= date('2017-01-01'))

df_top_tags_month %>% head()               
```

## Monthly by Tags

```{r}
plot <- ggplot(df_top_tags_month, aes(x=month_posted, y=count, fill=tag)) +
  geom_bar(alpha=0.75, stat="identity") +
  geom_smooth(se=F, method="lm", color="black", size=0.5) +
  scale_x_datetime(date_breaks='3 months', date_labels='%b') +
  scale_y_continuous(labels=comma) +
  facet_wrap(~ tag, nrow=10, ncol=4, scales="free_y") +
  labs(title='New Stack Overflow Questions for the Top 40 Tags',
       subtitle='From January 2017 to November 2017',
       x='Month Question Was Posted',
       y='# of Questions Posted With Tag') +
  theme(legend.position = 'none',
        plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')))

ggsave('monthly_count_tag.png', plot, width=6, height=8)
```

## Day/Hour of Week tags were posted

Related: https://stackoverflow.blog/2017/04/19/programming-languages-used-late-night/

```{r}
df_top_tags_hr_doy <- df %>%
  select(hour_posted, weekday_posted, tags) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
  group_by(hour_posted, weekday_posted, tag) %>%
  summarize(count=n()) %>%
  ungroup() %>%
  group_by(tag) %>%
  mutate(proportion=count/sum(count))

df_top_tags_hr_doy %>% head() 
```

```{r}
plot <- ggplot(df_top_tags_hr_doy, aes(x=hour_posted, y=fct_rev(weekday_posted), fill=proportion)) +
  geom_raster(stat="identity", interpolate=F) +
  geom_vline(xintercept=9, color="white", size=0.5, alpha=0.75) +
  geom_vline(xintercept=17, color="white", size=0.5, alpha=0.75) +
  scale_x_discrete() +
  scale_y_discrete() +
  scale_fill_viridis(option="inferno", labels=percent) +
  facet_wrap(~ tag, nrow=10, ncol=4) +
  labs(title='New Stack Overflow Questions for the Top 40 Tags, by Time Posted',
       subtitle='From January 2017 to November 2017, Vertical Lines Indicate 9 AM - 5 PM Eastern',
       x='Hour Question Was Posted (12 AM - 11 PM Eastern Time)',
       y='Day of Week Question Was Posted',
       fill='Proportion of All Questions\nPosted w/ Tag',
       caption = "Max Woolf — minimaxir.com") +
  theme(legend.position = 'top',
        plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
        axis.text.y = element_text(size = 5),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6),
        legend.key.width = unit(1, unit='cm'),
        legend.key.height = unit(0.25, unit='cm'),
        legend.margin = margin(c(0, 0, -0.5, 0), unit='cm'))

ggsave('monthly_count_hr_doy.png', plot, width=6, height=8)
```

## Day/Week To Receive Answer

```{r}
df_top_tags_hr_doy_answer <- df %>%
  select(hour_posted, weekday_posted, tags, f_answer_in, a_answer_in) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
  group_by(hour_posted, weekday_posted, tag) %>%
  summarize(count=n(), perc_f_answer = sum(f_answer_in)/count, perc_a_answer = sum(a_answer_in)/count) %>%
  ungroup()

df_top_tags_hr_doy_answer %>% head() 
```

## Tag Distributions

```{r}
df_top_tags_distribution <- df %>%
  select(month_posted, tags, view_count, time_to_f, time_to_a) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag)), month_posted >= date('2017-01-01'))

df_top_tags_distribution %>% head()
```

```{r}
plot <- ggplot(df_top_tags_distribution, aes(x=month_posted, y=view_count, fill=tag, group=month_posted)) +
          geom_boxplot() +
          scale_x_datetime(date_breaks='3 months', date_labels='%b') +
          scale_y_continuous(labels=comma) +
          facet_wrap(~ tag, nrow=10, ncol=4, scales="free_y") +
          labs(title='New Stack Overflow Questions for the Top 40 Tags',
                subtitle='From January 2017 to November 2017',
               x='Month Question Was Posted',
               y='# of Questions Posted With Tag') +
          theme(legend.position = 'none',
                plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')))

ggsave('monthly_views_boxplot_tag.png', plot, width=6, height=8)
```

## Tag Wordcloud

```{r}
df_top_tags_words <- df %>%
                select(tags, title) %>%
                unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
                filter(tag %in% (df_top_tags %>% head(20) %>% pull(tag))) %>%
                unnest_tokens(word, title, token="words") %>%
                filter(!(word %in% stop_words$word)) %>%
                group_by(tag, word) %>%
                summarize(count=n()) %>%
                ungroup() %>%
                group_by(tag) %>%
                mutate(max_norm = count/max(count)) %>%
                arrange(desc(count)) %>%
                top_n(20) %>%
                arrange(tag, desc(count))

df_top_tags_words %>% head(50) 
```

Wordcloud trick in ggplot2 adapted from [Mhairi Mcneill's blog post](http://mhairihmcneill.com/blog/2016/04/05/wordclouds-in-ggplot.html).

```{r}
set.seed(123)   # For geom_text_repel

plot <- ggplot(df_top_tags_words, aes(x = 1, y = 1, size = max_norm, label = word, color=max_norm)) +
  geom_text_repel(segment.size = 0, force = 100, family="Roboto Condensed Bold") +
  scale_size(range = c(2, 4), guide = FALSE) +
  scale_color_viridis(end=0.8, discrete=F, option="inferno", guide = FALSE) +
  scale_y_continuous(breaks = NULL) +
  scale_x_continuous(breaks = NULL) +
  facet_wrap(~ tag, nrow=10, ncol=4) +
  labs(title='New Stack Overflow Questions for the Top 40 Tags',
       subtitle='From January 2017 to November 2017, Sorted by Lowest Median Time to Accepted Answer',
       x='Month Question Was Posted',
       y='# of Questions Posted With Tag') +
  theme(legend.position = 'none',
        plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank()
  )

ggsave('so_tag_wordcloud.png', plot, width=6, height=8)
```

## Distribution of Times to an acceptable answer

```{r}
df_tag_median <- df_top_tags_distribution %>%
  group_by(tag) %>%
  summarize(med = median(time_to_a, na.rm=T)) %>%
  arrange(med)

df_tag_median
```


```{r}
plot <- ggplot(df_top_tags_distribution %>% filter(!is.na(time_to_a)) %>% mutate(tag = fct_relevel(factor(tag), (df_tag_median %>% pull(tag)))), aes(x=time_to_a, fill=tag)) +
  geom_histogram() +
  geom_vline(data=df_tag_median %>% mutate(tag = fct_relevel(factor(tag), (df_tag_median %>% pull(tag)))), aes(xintercept=med), linetype="dashed") +
  scale_x_continuous(limits = c(0,4*60*60), labels=0:4, breaks=seq(0,4*60*60, 60*60)) +
  #scale_x_log10(limits = c(1,10^3*60*60), labels=c(0, 10^(0:3)), breaks=c(0, 10^(0:3)*60*60)) +
  scale_y_continuous(labels=comma) +
  facet_wrap(~ tag, nrow=10, ncol=4, scales="free_y") +
  labs(title='New Stack Overflow Questions for the Top 40 Tags',
       subtitle='From January 2017 to November 2017, Sorted by Lowest Median Time to Accepted Answer',
       x='Month Question Was Posted',
       y='# of Questions Posted With Tag') +
  theme(legend.position = 'none',
        plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')))

ggsave('acceptable_answer_density.png', plot, width=6, height=8)
```


## Barchart of Acceptable Language Percents

```{r}
df_top_tags_distribution <- df %>%
  select(month_posted, tags, view_count, time_to_f, time_to_a) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  filter(tag %in% (df_top_tags %>% pull(tag)), month_posted >= date('2017-01-01')) %>%
  mutate(is_answered = ifelse(!is.na(time_to_a), 1, 0))

df_top_tags_distribution %>% head()
```


```{r}
df_acceptable_percs <- df_top_tags_distribution %>%
  group_by(tag) %>%
  summarize(count=n(),
            med_time_to_f = median(time_to_f, na.rm=T),
            med_time_to_a = median(time_to_a, na.rm=T),
            perc_is_answered = sum(is_answered) / count
  ) %>%
  arrange(desc(perc_is_answered))

df_acceptable_percs %>% head()
```

### Best/Worst Languages

This doesn't necessairly imply one tool is "better" than another, the difference may be due to question difficulty and the number of people skilled in the technology.

```{r}
df_acceptable_percs_subset <- df_acceptable_percs %>%
  head(30) %>%
  mutate(tag = fct_rev(as_factor(tag)))

plot <- ggplot(df_acceptable_percs_subset, aes(x=tag, y = perc_is_answered, fill=perc_is_answered)) +
  geom_bar(stat='identity') +
  geom_text(aes(label=percent(perc_is_answered), color=perc_is_answered), hjust=-0.25, family="Source Sans Pro Bold", size=3.5) +
  coord_flip() +
  scale_color_viridis(option='inferno', limits=c(0,1), guide=F) +
  scale_fill_viridis(option='inferno', limits=c(0,1), guide=F) +
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(title='Top Tags on Stack Overflow for Questions w/ Accepted Answers',
       subtitle='For Questions Asked From January 2017 to November 2017, out of Top 1,000 Tags',
       x='Stack Overflow Question Tag',
       y='% of Tagged Questions Which Have an Accepted Answer',
       caption = "Max Woolf — minimaxir.com") +
  theme(legend.position = 'none',
        plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
        axis.title.y=element_blank())

ggsave('acceptable_answer_top_30.png', plot, width=6, height=6)
```

```{r}
df_acceptable_percs_subset <- df_acceptable_percs %>%
  tail(30) %>%
  mutate(tag = fct_rev(as_factor(tag)))

plot <- ggplot(df_acceptable_percs_subset, aes(x=tag, y = perc_is_answered, fill=perc_is_answered)) +
  geom_bar(stat='identity') +
  geom_text(aes(label=percent(perc_is_answered), color=perc_is_answered), hjust=-0.25, family="Source Sans Pro Bold", size=3.5) +
  coord_flip() +
  scale_color_viridis(option='inferno', limits=c(0,1), guide=F) +
  scale_fill_viridis(option='inferno', limits=c(0,1), guide=F) +
  scale_y_continuous(labels=percent, limits=c(0,1)) +
  labs(title='Bottom Tags on Stack Overflow for Questions w/ Accepted Answers',
       subtitle='For Questions Asked from Janu
ary 2017 to November 2017, out of Top 1,000 Tags',
x='Stack Overflow Question Tag',
y='% of Tagged Questions Which Have an Accepted Answer',
caption = "Max Woolf — minimaxir.com") +
  theme(legend.position = 'none',
        plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
        axis.title.y=element_blank())

ggsave('acceptable_answer_bottom_30.png', plot, width=6, height=6)
```

# Adjacency Matrix for Top Tags

```{r}
df_tag_adjacency <- df %>%
  select(id, tags, time_to_a) %>%
  unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
  filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))) %>%
  inner_join(
    df %>%
      select(id, tags) %>%
      unnest_tokens(tag, tags, token=stringr::str_split, pattern=fixed("|")) %>%
      filter(tag %in% (df_top_tags %>% head(40) %>% pull(tag))), by=c("id" = "id")) %>%
  filter(tag.x != tag.y) %>%
  mutate(is_answered = ifelse(!is.na(time_to_a), 1, 0)) %>%
  group_by(tag.x, tag.y) %>%
  summarize(count=n(),
            total_answered = sum(is_answered),
            perc_answered = total_answered/count) %>%
  arrange(tag.x, desc(count))
```

```{r}
plot <- ggplot(df_tag_adjacency %>% filter(count >= 1000), aes(x=tag.x, y = tag.y, fill=count)) +
  geom_raster() +
  scale_fill_viridis(option='inferno', trans='log10', breaks=10^(2:5), labels=comma) +
  labs(title='Top Tags on Stack Overflow for Questions w/ Accepted Answers',
       subtitle='For Questions Asked From January 2017 to November 2017, out of Top 1,000 Tags',
       x='Stack Overflow Question Tag',
       y='% of Tagged Questions Which Have an Accepted Answer',
       caption = "Max Woolf — minimaxir.com") +
  theme(
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.x = element_text(angle = 270, hjust = 0, vjust = 0.5),
    legend.position = 'top',
    plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
    #axis.text.y = element_text(size = 5),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 6),
    legend.key.width = unit(1, unit='cm'),
    legend.key.height = unit(0.25, unit='cm'),
    legend.margin = margin(c(0, 0, -0.1, 0), unit='cm'))

ggsave('so_tag_adjacency.png', plot, width=6, height=6)
```

```{r}
plot <- ggplot(df_tag_adjacency %>% filter(count >= 1000), aes(x=tag.x, y = tag.y, fill=perc_answered)) +
  geom_raster() +
  scale_fill_viridis(option='inferno', limits=c(0,1), breaks=pretty_breaks(4), labels=percent) +
  labs(title='Top Tags on Stack Overflow for Questions w/ Accepted Answers',
       subtitle='For Questions Asked From January 2017 to November 2017, out of Top 1,000 Tags',
       x='Stack Overflow Question Tag',
       y='% of Tagged Questions Which Have an Accepted Answer',
       caption = "Max Woolf — minimaxir.com") +
  theme(
    axis.title.y=element_blank(),
    axis.title.x=element_blank(),
    axis.text.x = element_text(angle = 270, hjust=0, vjust = 0.5),
    legend.position = 'top',
    plot.title = element_text(size=10, family="Source Sans Pro Bold", margin=margin(t = -0.1, b = 0.0, unit='cm')),
    #axis.text.y = element_text(size = 5),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 6),
    legend.key.width = unit(1, unit='cm'),
    legend.key.height = unit(0.25, unit='cm'),
    legend.margin = margin(c(0, 0, -0.1, 0), unit='cm'))

ggsave('so_tag_adjacency_percent.png', plot, width=6, height=6)
```

# LICENSE

The MIT License (MIT)

Copyright (c) 2017 Max Woolf

Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the "Software"), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.